Temporal stability in human interaction networks

### 2015-06-30 paper.tex
adicionadas 2 paginas incorporadas e anotadas da revisao do ricardo

\begin{abstract}
In this study, we demonstrate a remarkably stable activity in human interaction networks. The activity along time and topology evolution were investigated in four e-mail lists by considering window sizes from 50 to 10,000 messages, which were made to slide and generate snapshots of the network in a timeline. Furthermore, the activity in terms of seconds, minutes, hours, days and months, is practically the same for all lists. The activity of participants followed the expected scale-free behavior, thus allowing us to establish three classes of vertices by comparing with the Erd\"os-R\'enyi model, namely hubs, intermediary and peripheral vertices. The relative size of these three sectors did not vary with time and was essentially the same for all e-mail lists. Typically, 3-12\% of the vertices are hubs, 15-45\% are intermediary and the remainder are peripheral vertices. The metrics that contribute most to the dispersion of participants in the topological measures space were centrality measurements (degree, strength and betweenness), followed by symmetry-related metrics and then clustering coefficient. Similar results for the distribution of participants in the three categories and for the relative importance of the topological metrics were obtained for 12 additional networks from Facebook, Twitter and Participa.br. Consistent with expectations from the literature, these properties may be general for human interaction networks, which has important implications in establishing a typology based on objective, quantitative criteria.
\end{abstract}

@@

\subsection{Related work}\label{sec:related}
Research on network evolution often considers solely network growth, in which there is a monotonic increase in the number of events~\cite{barabasiEvo}. Exceptions are reported in this section, with emphasis on those more closely related to the present article.

@@

Unreciprocated edges often exceed 50\% in the networks analyzed, which matches empirical evidence from the literature~\cite{newmanEvolving} and motivated the inclusion of symmetry metrics in our analysis. No correlation of topological characteristics and geographical coordinates was found~\cite{barabasiGeo}, therefore geographical positions were not considered in our study. Gender related behavior in mobile phone datasets was indeed reported~\cite{barabasiSex}, but this was not considered in the present article because email messages and addresses have no gender related metadata~\cite{gmanePack}.

@@

We analyzed many email lists (and data from Twitter, Facebook and Participa.br) but selected only four in order to make a thorough analysis, from which general properties can be inferred. These lists, selected as representing both a diverse set and ordinary lists, are:
\begin{itemize}
	\item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in GMANE.}, with participants holding hybrid artistic and technological interests, from different countries. English is the language used the most. Abbreviated as LAU from now on.
	\item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in GMANE.}, with participants from different countries, and English is the language used the most. A more technical and less active version of LAU. Abbreviated LAD from now on.
	\item Development list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in GMANE.}, with computer programmers from different countries. English is the language used the most. Abbreviated as CPP from now on.
	\item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in GMANE.}, with Brazilian activists holding digital culture interests. Portuguese is the most used language, although Spanish and English are also incident. Abbreviated MET from now on.
\end{itemize}

The first 20,000 messages of each list were considered, with total timespan, authors, threads and missing messages indicated in Table~\ref{tab:genLists}.
In subsidiary experiments we considered 140 additional email lists, also retrieved from the GMANE public database, to analyze the interdependence between the number of participants and the number of threads. Furthermore, we used 12 additional networks from Facebook (8), Twitter (2) and Participa.br (2) to grasp the generality of the results driven from email lists.

\begin{table}
	\centering
	\caption{Columns $date_1$ and $date_M$ have dates of first and last messages from the 20,000 messages considered in each email list.
		$N$ is the number of participants (number of different email addresses).
		$\Gamma$ is the number of threads (count of messages without antecedent).
		$\overline{M}$ is the number of messages missing in the 20,000 collection, $100\frac{23}{20000}=0.115$ percent in the worst case.
	}
	\label{tab:genLists}
	\begin{tabular}{|l|c|c|c|c|c|}\hline
		list & $date_1$ & $date_{M}$    & $N$  & $\Gamma$ & $\overline{M}$ \\\hline
		\input{tables/tab1Geral}
	\end{tabular}
\end{table}

\section{Characterization methods}\label{sec:carac}
The email lists and the networks generated from them were characterized using five procedures, namely: 1) statistics of activity along time, in scales from seconds to years; 2) sectioning of the networks in hubs, intermediary and peripheral vertices; 3) dispersion of basic topological metrics; 4) iterative visualization and data inspection.
Each of these procedures are described below.

\subsection{Time activity statistics}\label{sec:mtime}
Messages were counted along time with respect to seconds, minutes, hours, days of the week, days of the month, and months of the year. This resulted in histograms from which patterns could be drawn. The ratio $\frac{b_h}{b_l}$ between the highest and lowest incidences on the histograms served as a hint of how close the observed distribution is to a uniform distribution.

The average and the dispersion were taken using circular statistics, in which each $measurement$ (data point) is represented as a complex number with modulus equal to one, $z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, where $\theta=measurement\frac{2\pi}{period}$. The moments $m_n$, lengths of moments $R_n$, mean angle $\theta_\mu$, and rescaled mean angle $\theta_\mu'$ are defined as:

>>>>

\begin{abstract}
In this study, we demonstrate a remarkably stable activity in human interaction networks. The activity along time and topology evolution was investigated in four e-mail lists by considering window sizes from 50 to 10,000 messages, which were made to slide and generate snapshots of the network in a timeline. Notably, the activity in terms of seconds, minutes, hours, days and months, is practically the same for all lists. The activity of participants followed the expected scale-free behavior, thus allowing us to establish three classes of vertices by comparing against the Erd\"os-R\'enyi model, namely hubs, intermediary and peripheral vertices. The relative size of these three sectors did not vary with time and was essentially the same for all e-mail lists. Typically, 3-12\% of the vertices are hubs, 15-45\% are intermediary and the remainder are peripheral vertices. The metrics that contribute most to the dispersion of participants in the topological measures space were centrality measurements (degree, strength and betweenness), followed by symmetry-related metrics and then clustering coefficient. Similar results for the distribution of participants in the three categories and for the relative importance of the topological metrics were obtained for 12 additional networks from Facebook, Twitter and Participa.br. Consistent with expectations from the literature, these properties may be general for human interaction networks, which has important implications in establishing a typology based on objective, quantitative criteria.
\end{abstract}

@@

\subsection{Related work}\label{sec:related}
Research on network evolution is often restricted to network growth, in which there is a monotonic increase in the number of events~\cite{barabasiEvo}. Exceptions are reported in this section, with emphasis on those more closely related to the present article.

@@

Unreciprocated edges often exceed 50\% in the analyzed networks, which matches empirical evidence from the literature~\cite{newmanEvolving} and motivated the inclusion of symmetry metrics in our analysis. No correlation of topological characteristics and geographical coordinates was found~\cite{barabasiGeo}, therefore geographical positions were not considered in our study. Gender related behavior in mobile phone datasets was indeed reported~\cite{barabasiSex}, but this was not considered in the present article because email messages and addresses have no gender related metadata~\cite{gmanePack}.

@@

We analyzed many email lists (and data from Twitter, Facebook and Participa.br) but selected only four in order to make a thorough analysis, from which general properties can be inferred. These lists, selected as representative of both a diverse set and ordinary lists, are:
\begin{itemize}
	\item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in GMANE.}, with participants holding hybrid artistic and technological interests, from different countries. English is the language used the most. Abbreviated as LAU from now on.
	\item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in GMANE.}, with participants from different countries, and English is the language used the most. A more technical and less active version of LAU. Abbreviated LAD from now on.
	\item Development list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in GMANE.}, with computer programmers from different countries. English is the language used the most. Abbreviated as CPP from now on.
	\item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in GMANE.}, with Brazilian activists holding digital culture interests. Portuguese is the most used language, although Spanish and English are also incident. Abbreviated MET from now on.
\end{itemize}

The first 20,000 messages of each list were considered, with total timespan, authors, threads and missing messages indicated in Table~\ref{tab:genLists}.
In subsidiary experiments we considered 140 additional email lists, also retrieved from the GMANE public database, to analyze the interdependence between the number of participants and the number of discussion threads. Furthermore, we used 12 additional networks from Facebook (8), Twitter (2) and Participa.br (2) to grasp the generality of the results driven from email lists.

\begin{table}
	\centering
	\caption{Columns $date_1$ and $date_M$ have dates of first and last messages from the 20,000 messages considered in each email list.
		$N$ is the number of participants (number of different email addresses).
		$\Gamma$ is the number of discussion threads (count of messages without antecedent).
		$\overline{M}$ is the number of messages missing in the 20,000 collection, $100\frac{23}{20000}=0.115$ percent in the worst case.
	}
	\label{tab:genLists}
	\begin{tabular}{|l|c|c|c|c|c|}\hline
		list & $date_1$ & $date_{M}$    & $N$  & $\Gamma$ & $\overline{M}$ \\\hline
		\input{tables/tab1Geral}
	\end{tabular}
\end{table}

\section{Methods}\label{sec:carac}
The email lists and the networks generated from them were characterized using five procedures, namely: 1) statistics of activity along time, in scales from seconds to years; 2) sectioning of the networks in hubs, intermediary and peripheral vertices; 3) dispersion of basic topological metrics; 4) iterative visualization and data inspection.
Each of these procedures are described below.

\subsection{Time activity statistics}\label{sec:mtime}
Messages were counted over time with respect to seconds, minutes, hours, days of the week, days of the month, and months of the year. This resulted in histograms from which patterns could be drawn. The ratio $\frac{b_h}{b_l}$ between the highest and lowest incidences on the histograms served as a clue of how close the observed distribution is to a uniform distribution.

The average and the dispersion were taken using circular statistics, in which each $measurement$ (data point) is represented as a unit complex number, $z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, where $\theta=measurement\frac{2\pi}{period}$. The moments $m_n$, lengths of moments $R_n$, mean angle $\theta_\mu$, and rescaled mean angle $\theta_\mu'$ are defined as:
